### 2024-07-09 Article/Article-english.tex
write up

@@

>>>>

We denote a~\textit{$\mathbb{Z}$-linear span} of~$\alpha_1$, $\ldots$, $\alpha_n$
by~$\mathrm{Span}_\mathbb{Z}(\alpha_1, \ldots, \alpha_n) = \{ k_1 \alpha_1 + \ldots + k_n \alpha_n \ | \ k_1,\allowbreak \ldots,\allowbreak k_n \in \mathbb{Z} \}$.
We say that $\alpha_1$, $\ldots$, $\alpha_n$ span the~whole lattice $\mathbb{Z}^k$ if
$\mathrm{Span}_\mathbb{Z}(\alpha_1, \ldots, \alpha_n) = \mathbb{Z}^k$.

@@

Now we will show that the~obtained necessary condition is also sufficient for~the~group $G = {\mathbb{C}^\times}$.

We will use the~existence of~\textit{Smith's normal form} for~integer matrices \cite{Smth60}.
Let $\mathrm{diag}^{m \times n}_r(x_1,\allowbreak \ldots,\allowbreak x_r)$ be a~diagonal matrix $\mathrm{diag}(x_1,\allowbreak \ldots,\allowbreak x_r)$
augmented (or cut off) from the~bottom right by~zeroes to a~matrix of size $m \times n$.

\medskip\noindent\textbf{Theorem} (on~the~existence of~Smith's normal form).\emph{
    $$
        \forall \alpha \in \mathbb{Z}^{m \times n}{:}\
        \exists \beta_1 \in \mathrm{GL}^m(\mathbb{Z}), \beta_2 \in \mathrm{GL}^n(\mathbb{Z}){:} \
        \exists \varepsilon_1, \ldots, \varepsilon_r \in \mathbb{Z} \setminus \{0\}{:} \
        \beta_1 \alpha \beta_2 = \mathrm{diag}^{m \times n}_r(\varepsilon_1, \ldots, \varepsilon_r),
    $$
    where $\varepsilon_1 \mid \varepsilon_2 \mid \ldots \mid \varepsilon_r$ (the so~called invariant factors) and~$r = \mathrm{rank}(\alpha)$.
}\medskip

Since the numbers $\varepsilon_1$, $\ldots$, $\varepsilon_r$ are chosen in~a~unique way up to~the~invertible element,
that is, up to $\pm 1$ in~case of~$\mathbb{Z}$, we can always choose them positive. For them we denote
the~matrix $\mathrm{diag}^{m \times n}_r(\varepsilon_1, \ldots, \varepsilon_r)$ by~$\mathrm{SNF}(\alpha)$.

In~addition, since $\phi$ defines the~functor, it is clear that $\phi_\beta$ is bijective if $\beta \in \mathrm{GL}^n(\mathbb{Z})$.
Indeed, $\phi_\beta \circ \phi_{\beta^{-1}} = \phi_{\beta \beta^{-1}} = \phi_E = 1_{(K^\times)^n}$
and~$\phi_{\beta^{-1}} \circ \phi_\beta = \phi_{\beta^{-1} \beta} = \phi_E = 1_{(K^\times)^n}$. Therefore,
the~following lemma holds.

\medskip\noindent\textbf{Lemma~4.}\emph{
    $\phi_\alpha : (\mathbb{C}^\times)^k \rightarrow (\mathbb{C}^\times)^n$ is injective if and only if
    $$
        \mathrm{SNF}(\alpha) = \mathrm{diag}^{n \times k}_k(1, 1, \ldots, 1).
    $$
}

\noindent{\it Proof.}
    Using the~theorem, let us take unimodular matrices $\beta_1 \in \mathrm{GL}^n(\mathbb{Z})$ and~$\beta_2 \in \mathrm{GL}^k(\mathbb{Z})$
    such that $\beta_1 \alpha \beta_2 = \mathrm{SNF}(\alpha)$.
    Then, $\alpha = \beta_1^{-1} \mathrm{SNF}(\alpha) \beta_2^{-1}$ and~$\phi_\alpha = \phi_{\beta_1^{-1} \mathrm{SNF}(\alpha) \beta_2^{-1}}
                                                                                    = \phi_{\beta_1^{-1}} \circ \phi_{\mathrm{SNF}(\alpha)} \circ \phi_{\beta_2^{-1}}$.

    Since, by~the~remark above, $\phi_{\beta_1^{-1}}$ and~$\phi_{\beta_2^{-1}}$ are bijective, $\phi_\alpha$ is injective if and only if
    $\phi_{\mathrm{SNF(\alpha)}}$ is injective.

    $\phi_{\mathrm{SNF(\alpha)}}(t_1, \ldots, t_k) = (t_1^{\varepsilon_1}, \ldots, t_r^{\varepsilon_r}, 1, \ldots, 1)$, but $t \mapsto t^\varepsilon$
    is injective over $\mathbb{C}^\times$ only if $\varepsilon = \pm 1$ (otherwise $t^\varepsilon = t^\varepsilon \cdot 1 = t^\varepsilon \cdot u^\varepsilon = (tu)^\varepsilon$
    where $u \neq 1$ is~a~nontrivial $\varepsilon$-th root of~unity).
    By~choice of~signs, $\phi_{\mathrm{SNF(\alpha)}}$ is injective only if $\varepsilon_1 = \ldots = \varepsilon_r = 1$
    and~$r = k \leq n$.
\hfill$\Box$\medskip

Let us formulate one more auxiliary theorem \cite{TsikhSad14}.

\medskip\noindent\textbf{Theorem~1.}\emph{
    Let $\alpha \in \mathbb{Z}^{n \times k}$. The~following statements are equivalent:
    \begin{enumerate}
        \item The~rows of~the~matrix $\alpha$ span the~whole lattice $\mathbb{Z}^k$.
        \item The~minors of~the~maximum order of~the~matrix $\alpha$ are coprime.
        \item $\mathrm{SNF}(\alpha) = \mathrm{diag}^{n \times k}_k(1, \ldots, 1)$.
    \end{enumerate}
}

Thus, from lemma~4 and~theorem~1 we obtain a~necessary and~sufficient condition for~the~injectivity of~$\phi_\alpha$.

\medskip\noindent\textbf{Theorem.}\emph{
    $\phi_\alpha$ is injective over the~field $K = \mathbb{C}$ if and only if $\alpha_i$ span the~whole lattice $\mathbb{Z}^k$.
}

\section*{4. Monomial parameterizability of~torus subgroups}

As in~the~case of~the~theory of~curves and surfaces, we say that a~subgroup of~an~algebraic torus
is \textit{parameterizable} if it can be represented as an~image of~some mapping. Similarly, a~subgroup
is \textit{monomially parameterizable} if it can be represented as an~image of~$\phi_\alpha$ for some matrix $\alpha$.
